Appendix 1.B: Analysing global trends in within-country inequality

### 2023-01-22 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Moving code to shinylive app

@@

>>>>

from shiny import *
from plotnine import *
from pyodide.http import open_url
import pandas as pd
import numpy as np

#  ---------------------------------
#  ---------- Data prep functions -------------
#  ----------------------------------

# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

# Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')

# Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')

  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

#  ---------------------------------
#  ---------- Specifications -------------
#  ----------------------------------

# These parameters are not controlled in the app.
reference_col = 'Year'
group_by_col = 'Entity'

source_count_requirement = 2

#  ---------------------------------
#  ---------- App code -------------
#  ----------------------------------

def panel_box(*args, **kwargs):
    return ui.div(
        ui.div(*args, class_="card-body"),
        **kwargs,
        class_="card mb-3",
    )

app_ui = ui.page_fluid(
    {"class": "p-4"},
    ui.row(
        ui.column(
            4,
            panel_box(
                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
                ui.input_numeric(
                            "max_dist_from_refs",
                            "=/- (years)",
                            4,
                        ),
                ui.input_numeric(
                            "min_dist_between",
                            "Min distance between observations (years)",
                            23,
                        ),
                ui.input_action_button(
                    "download", "Download data and image", class_="btn-primary w-100"
                ),
            ),
            ui.navset_tab_card(
                ui.nav(
                    "Compare sources",
                    ui.input_select(
                "metric",
                "Metric",
                {
                    "Gini": "Gini",
                    "Top_10_share": "Top 10p.c. share",
                },
            ),
            ),
                ui.nav(
                  "Compare metrics"),
            ),
        ),
        ui.column(
            8,
            ui.output_plot("example_plot")
            ),
    ),
)

      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'

      df_pip = pd.read_csv(open_url(url))

      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'

      df_wid = pd.read_csv(open_url(url))

      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'

      df_regions = pd.read_csv(open_url(url))

      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'

      df_pop = pd.read_csv(open_url(url))

      def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

        pip_matches = pip_welfare_routine(
          df = df_pip,
          reference_vals = reference_vals,
          max_dist_from_refs = max_dist_from_refs,
          min_dist_between = min_dist_between,
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = value_col
        )

        return ref_pairs

        plot = (ggplot(plot_data
        , aes(x_axis, y_axis, alpha=0.5))
        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
        + facet_wrap('~ source')
        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
        + theme_light()
        + theme(legend_position='none')
        )

        # ----- Return outputs ------
        plot_and_tables = {
          "table_data": plot_data,
          "plot": plot,
          "summary": df_summary
        }

      plot_and_tables = prep_plot_and_tables(
          reference_vals = list(input.reference_vals()), 
          max_dist_from_refs = input.max_dist_from_refs(),
          min_dist_between = input.min_dist_between(),
          reference_col = reference_col,
          group_by_col = group_by_col,
          value_col = input.metric(),
          tolerance = tolerance,
          outlier_cut_off_upper = outlier_cut_off_upper,
          source_count_requirement = source_count_requirement
          )

      plot = plot_and_tables['plot']

@@

      # #  -----------------------------------
      # #  ------- Data prep functions -------
      # #  -----------------------------------

      # #  Function get matching for ref years
      # def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

      #     df = df.loc[:, [reference_col, group_by_col, value_col]]

      #     # Drop NAs
      #     df = df.dropna()

      #     # Calculate absolute distance from reference value
      #     df['ref_diff'] = abs(df[reference_col] - reference_val)

      #     # Drop any rows with a distance beyond threshold
      #     if not pd.isna(max_dist_from_ref):
      #       df = df.loc[df['ref_diff'] <= max_dist_from_ref]

      #     # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
      #     df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

      #     # Settle tie-breaks
      #     if tie_break == 'below':
      #       df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

      #     elif tie_break == 'above':
      #       df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

      #     df = df.drop('ref_diff', axis=1)

      #     return df

      #   def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

      #     # Make sure the pair of reference values are in ascending order
      #     reference_vals.sort()

      #     # Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

      #     # Find matches for lower reference value
      #     lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs[0], reference_col, group_by_col, value_col, 'below')

      #     # Find matches for higher reference value
      #     higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs[1], reference_col, group_by_col, value_col, 'above')

      #     # Merge the two sets of matches
      #     merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

      #     # Drop obs that do not have data for both ref values
      #     merged_df = merged_df.dropna()

      #     # Drop obs where the matched data does not meet the min distance requirement
      #     if not pd.isna(min_dist_between):

      #       # Store the names of the reference column returned from the two matches
      #       ref_var_high = f'{reference_col}{reference_vals[1]}'
      #       ref_var_low = f'{reference_col}{reference_vals[0]}'

      #      # Keep only rows >= to the min distance
      #       merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

      #     return merged_df

      #   def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

      #     # Specify the name of the column in which the income/consumption welfare definition is stored
      #     welfare_colname = 'welfare_type'

      #     # Creat dataframes for thee scenarios:
      #     # Scenario 1: only allow income data
      #     df_inc_filter = df.loc[df[welfare_colname] == "income", :]
      #     df_inc_filter.name = "Income"

      #     # Scenario 2: only allow consumption data
      #     df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
      #     df_cons_filter.name = "Consumption"
      #     # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
      #     df_mixed = df.copy()

      #     df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

      #     df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

      #     df_mixed.name = "Mixed"
      #     #  Store the scneario dataframes in a list
      #     df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

      #     # Run the matching function on each scenario
      #     scenario_matches = [merge_two_ref_matches(
      #       df_scenario, 
      #       reference_vals, 
      #       max_dist_from_refs, 
      #       min_dist_between, 
      #       reference_col, 
      #       group_by_col, 
      #       value_col) for df_scenario in df_scenarios]

      #     # Combine the first two scenarios.
      #     df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

      #     # Tidy up indexes
      #     df_combined_matches = df_combined_matches.reset_index()

      #     df_combined_matches = df_combined_matches.drop('level_1', axis=1)

      #     df_combined_matches = df_combined_matches\
      #         .rename(columns={"level_0": "pip_welfare"})

      #     # Add in third scenario.
      #     df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

      #     # add scenario name to te pip_welfare column
      #     df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

      #     # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
      #     # First count the matches
      #     df_combined_matches['match_count'] =        df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
      #     # Then drop any matches from the lowest priority where there are multiple matches
      #     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
      #     #  Repeat at the next level of priority
      #     df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
      #     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

      #     # Drop the match count column
      #     df_combined_matches = df_combined_matches.drop('match_count', axis=1)

      #     return df_combined_matches
